---
title: "P8105_HW5_TEM2171"
author: "Teresa Moore"
date: "2023-11-15"
output: github_document
---
```{r}
library(tidyverse)
```

# PROBLEM 2

Create a tidy dataframe containing data from all participants, including the subject ID, arm, and observations over time:

```{r}
all_participants_df <- 
  tibble(
    files = list.files("data"),
    path = str_c("data/", files)
  ) %>% 
  mutate(data = map(path, read_csv)) %>% 
  unnest()

all_participants_tidy_df <- 
  all_participants_df %>% 
  mutate(
    files = str_replace(files, ".csv", ""),
    group = str_sub(files, 1, 3)
  ) %>% 
  pivot_longer(
    cols = starts_with("week_"),
    names_to = "week",
    values_to = "outcome",
    names_prefix = "week_"
  ) %>% 
  mutate(week = as.numeric(week)) %>% 
  select(group, subj = files, week, outcome)
```

Make a spaghetti plot showing observations on each subject over time, and comment on differences between groups:

```{r}
all_participants_tidy_df |> 
  ggplot(aes(x = week, y = outcome, group = subj, color = group)) + geom_path() + facet_grid(~group)
```

For the control arm, the outcomes on average stay the same while for the experimental arm, we see on average an increaase in the outcomes.

#PROBLEM 3

```{r}
n <- 30
sigma <- 5
alpha <- 0.05
num_simulations <- 5000
```

```{r}
simulation_t_test = function(mu) {
  simulation_data = tibble(
    x = rnorm(n, mean = mu, sd = sigma),
  )

output = simulation_data |>
  t.test() |>
  broom::tidy() |>
  select(estimate, p.value)

}
```

```{r}
simulation_results = expand_grid(
  true_mu = c(0, 1, 2, 3, 4, 5, 6),
  iter = 1:num_simulations
) |>
  mutate(
    estimate = map(true_mu, simulation_t_test)
  ) |>
  unnest(estimate)
```

Make a plot showing the proportion of times the null was rejected (the power of the test) on the y axis and the true value of μ on the x axis. Describe the association between effect size and power

```{r}
simulation_results |>
  group_by(true_mu) |>
  summarize(
    proportion = sum(p.value<alpha)/num_simulations
  ) |>
  ggplot(aes(x=true_mu, y=proportion)) + geom_line()
```
As the true value of mu increases, the proportion of times the null was rejected, also known as the power, also increases. There is a higher rate of increase in proportion between mu values 1 and 3, before it levels out to 100% at mu= 4. 

a plot showing the average estimate of μ̂ on the y axis and the true value of μ on the x axis. Make a second plot (or overlay on the first) the average estimate of μ̂ only in samples for which the null was rejected on the y axis and the true value of μ on the x axis: 

```{r}
all_samples_df = simulation_results |>
  group_by(true_mu) |>
  summarize(
    mean_mu = mean(estimate)
  )

rejected_null_df = simulation_results |>
  filter(p.value < alpha) |>
  group_by(true_mu) |>
  summarize(
    mean_mu = mean(estimate)
  )

ggplot() +
  geom_line(data=all_samples_df, aes(x=true_mu, y=mean_mu, color="All Samples"), show.legend=TRUE) +
  geom_line(data=rejected_null_df, aes(x=true_mu, y=mean_mu, color="Rejected Null"), show.legend=TRUE) +
  labs(color="Legend") +
  theme_minimal()
```

The sample average of μ̂ across tests for which the null is rejected is no approximately equal to the true value of μ for lower values of mu between 0-4. Above 4, it is approximately equal.
